Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-10-10 Manuscript/manuscript.tex
Updates and comments in methods section

Lipid bilayer structure without ions was validated against NMR and 
\mbox{x-ray} scattering experiments by calculating order parameters
for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
lipid molecules in simulations with atomic resolution, while the latter
validates the dimensions of the lipid bilayer (thickness and area per
molecule)~\cite{ollila16}.

The order parameters were calculated from simulations for all C-H bonds
in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
from equation
\begin{equation}
  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
\end{equation}
where $f_\alpha(q_z)$ is the density of atomic scattering length, 
$\rho_s$ is the density of solvent scattering length in the bulk region,
$n_\alpha (z)$ is the number density of atom $\alpha$ and
$z$ is the distance from the membrane centre along its normal 
spanning until the water bulk region, $D$.

@@

\subsubsection{Simulations of POPC bilayers in aqueous ions}
The simulated systems consisted of POPC bilayer in pure water
or in varying salt concentrations.
In particular, the periodic orthorhombic simulation box contained 
128 POPC molecules and approximately 50~water molecules per each lipid.
As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
which is currently the most accurate three site rigid water model. 
In order to test transferability of our newly developed ECC-lipids model, 
we also performed several additional simulations with 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
presented in Supporting Information (SI). 
We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

@@

\begin{table}
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular}
\end{table}

>>>>

The structure of lipid bilayers in simulations without ions were validated against NMR and 
\mbox{x-ray} scattering experiments by calculating order parameters
for C-H bonds and form factors. The former validates the structures sampled by individual
lipid molecules in simulations with atomic resolution, while the latter
validates the dimensions of the lipid bilayer (thickness and area per
molecule)~\cite{ollila16}.

The order parameters were calculated from simulations for all C-H bonds
in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
from equation \todo{As Hector suggested, it might be better to write the simpler
form for this equation.}
\begin{equation}
  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
\end{equation}
where $f_\alpha(q_z)$ is the density of atomic scattering length, 
$\rho_s$ is the density of solvent scattering length in the bulk region,
$n_\alpha (z)$ is the number density of atom $\alpha$ and
$z$ is the distance from the membrane centre along its normal 
spanning until the water bulk region, $D$.

@@

\subsubsection{Simulations of POPC bilayers in aqueous ions}
Simulations of POPC bilayer in pure water or in varying salt concentrations
contained 128 POPC molecules and approximately 50~water molecules per each lipid
in the periodic orthorhombic simulation box.
As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
which is currently the most accurate three site rigid water model
\todo{More justification for the choosing the OPC3 water model are needed.
It might be good to show the comparison with the scattering data in bulk water in SI.}. 
In order to test transferability of our newly developed ECC-lipids model, 
we also performed several additional simulations with 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}. 
%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
%this is done in Amber by default. If this is correct, the information in the
%table do not match with this sentence. Based on semi-extensive testing I made few years
%ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

@@

\begin{table}
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
  \todo{Result with normal TIP3P missing?}
\end{table}
